Project assignment - Tahoe-LAFS with SHA-3 candidates

### 2010-12-11 paper/main.tex
Added section on immutable file exclusion in Discussion chapter

For the same reasons, some of the built-in unit tests of Tahoe-LAFS fail when
changing this line, because they contain test fixtures made with the included
SHA-256.  An implication of this is that uploading of mutable files renders
impossible, since the internal verification fails. For this reason, \emph{only
immutable files} are tested and measured in this paper. Implications of this is
discussed further in Chapter \ref{ch:discussion}.

@@

\vspace{12pt}
\noindent For every \ac{SHA}-3 candidate:
        \begin{enumerate}
            \item Install SHA3lib on all nodes
            \item Replace used hash algorithm in Tahoe-LAFS with current
            candidate
            \item Install Tahoe-LAFS
            \item Create and start one Tahoe-LAFS introducer
            \item Create and start Tahoe-LAFS storage nodes on 3 nodes
            \item Create a Tahoe-LAFS client on a previously unnused node
            \item Do testing
            \item clean up
        \end{enumerate}

@@

\section{\ac{SUPERCOP} results}
We see in tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
\ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
\ref{tbl:supercop:8} that in general BLAKE, \ac{BMW}, Luffa, Shabal and Simd
outperform \ac{SHA}-256. However, given that Tahoe-LAFS uses double
\ac{SHA}-256, this means that it will also have to hash another 32 bytes for
every message it hashes. Since \ac{SUPERCOP} does not have benchmark for 32
byte messages, we will have to estimate the amount of cycles required for this.
If we assume that cycles/bytes is linear from 8 byte to 64 byte, we get the
following estimate for the speed of hashing 32 bytes:

@@

We can then estimate the result of each message by the following formula:
$$\frac{median * size + 118.98*32}{size} = SHA256d estimate$$

\input{tables/supercop_sha256d}
Estimates based on this calculation are shown in Table
\ref{tbl:supercop:sha256d}. What we can see is that for lower lengths of data,
more of the candidates are faster than double \ac{SHA}-256. However, given the
practical situation that Tahoe-\ac{LAFS} is a file system, most files processed
will have be larger than the file sizes estimated here. Even though not all
hashing operations done in Tahoe are related to the file size, we expect that
only BLAKE, \ac{BMW}, Luffa, Shabal and SIMD will be able to outperform double
\ac{SHA}-256.

\section{Validity of \ac{SUPERCOP} results}
Although we believe the \ac{SUPERCOP} toolkit and results give good indication
of how fast the different candidate implementations are, we do have some
comments on it.

\paragraph{Implementations does not conform to \ac{NIST} specification} for
many of the candidates. By this we mean that there is no Update()-function
available for all implementation, and that \ac{SUPERCOP} only measures speed of
hashing full byte strings. This is also a hassle for a programmer who want to
use the implementations in \ac{SUPERCOP}

\paragraph{Incorrect implementations} are present in \ac{SUPERCOP}. By this we
mean implementations that does not verify the \ac{KAT}s.

@@

% This chapter should contain (in listed and grouped order):
% -----------------------------------------------------------
% Describe inconsistencies in the numbers, and try to explain them.
% Time spent not hashing varies because: Variable use of CPU regs?
%
% Compare our results with general (SUPERCOP) results.
% Optimized candidates - which gain could be expected from more optimized
%   candidates?
% Mention that some SUPERCOP candidates are missing UPDATE-functionality.
%
% How much gain can Tahoe-LAFS get by switching to SHA-3?
% Time gain of removing ``double'' SHA-256. Double only in digest().
%
% Implications of only testing immutable files? (mutable files also
%   include directories)
% Adding a capability to support new hash functions in tahoe. (Our
%   current way breaks backwards compability)
%
% What if we had used another coding scheme? - does Cython lower the
%   candidates performance?
%
% More general stuff. What could have been done differently? Criticize
% ourselves.

@@

A somewhat short conclusion summing up what has been done, and results and
discussion

New SHA-3 capability in Tahoe-LAFS (GSOC ref) to get our code to trunk

>>>>

For the same reasons, some of the built-in unit tests of Tahoe-LAFS fail when
changing this line, because they contain test fixtures made with the included
SHA-256.  An implication of this is that uploading of mutable files renders
impossible, since the internal verification fails. For this reason, \emph{only
immutable files} are tested and measured in this paper. Implications of this are
discussed further in Chapter \ref{ch:discussion}.

@@

\vspace{12pt}
\noindent For every \ac{SHA}-3 candidate:
        \begin{enumerate}
            \item Install SHA3lib on all nodes
            \item Replace used hash algorithm in Tahoe-LAFS with current
            candidate
            \item Install Tahoe-LAFS
            \item Create and start one Tahoe-LAFS introducer
            \item Create and start Tahoe-LAFS storage nodes on 3 nodes
            \item Create a Tahoe-LAFS client on a previously unnused node
            \item Do testing
            \item Clean up
        \end{enumerate}

@@

\section{\ac{SUPERCOP} results}
% Time gain of removing ``double'' SHA-256. Double only in digest().

We see in Tables \ref{tbl:supercop:long}, \ref{tbl:supercop:4096},
\ref{tbl:supercop:1536}, \ref{tbl:supercop:576}, \ref{tbl:supercop:64} and
\ref{tbl:supercop:8} that in general BLAKE, \ac{BMW}, Luffa, Shabal and Simd
outperform \ac{SHA}-256. However, given that Tahoe-LAFS uses double
\ac{SHA}-256, this means that it will also have to hash another 32 bytes for
every message it returns the digest for. Since \ac{SUPERCOP} does not have
benchmark for 32 byte messages, we will have to estimate the amount of cycles
required for this.  If we assume that cycles/bytes is linear from 8 byte to 64
byte, we get the following estimate for the speed of hashing 32 bytes:

@@

We can then estimate the result of each message by the following formula:
$$\frac{median * size + 118.98*32}{size} = SHA256d$$ \emph{estimate}

\input{tables/supercop_sha256d}

Estimates based on this calculation are shown in Table
\ref{tbl:supercop:sha256d}. What we can see that for lower lengths of data,
more of the candidates are faster than double \ac{SHA}-256. However, given the
practical situation that Tahoe-\ac{LAFS} is a file system, most files processed
will be larger than the file sizes estimated here. Even though not all
hashing operations done in Tahoe are related to the file size, we expect that
only BLAKE, \ac{BMW}, Luffa, Shabal and SIMD will be able to outperform double
\ac{SHA}-256.

\section{Validity of \ac{SUPERCOP} results}
% Mention that some SUPERCOP candidates are missing UPDATE-functionality.

Although we believe the \ac{SUPERCOP} toolkit and results give good indication
of how fast the different candidate implementations are, we do have some
comments on it.

\paragraph{Implementations does not conform to \ac{NIST} specification} for
many of the candidates. By this we mean that there is no {\verb Update() }
function available for all implementations, and that \ac{SUPERCOP} only
measures speed of hashing full byte strings. This is also a hassle for a
programmer who wants to use the implementations in \ac{SUPERCOP}

\paragraph{Incorrect implementations} are present in \ac{SUPERCOP}. By this we
mean implementations that does not conform to the \ac{KAT}s supplied.

@@

\section{Exclusion of mutable file testing}
% Implications of only testing immutable files? (mutable files also
%   include directories)
% Adding a capability to support new hash functions in tahoe. (Our
%   current way breaks backwards compability)

Since our minor change in the Tahoe-\ac{LAFS} code broke internal checks, we
were not able to test mutable files, and hence not directories.

We argue that this has minor implications for our testing results. Both because
mutable files are not applicable for most use cases, since deleting an old
version and uploading a new file gives the same functionality, and because
immutable files uses less logic and therefore are more reliable and faster in
some cases. One should bear in mind that Tahoe-\ac{LAFS} is not a revision
control system, and mutable files are not intended for this purpose.

To be able to extend the Tahoe-LAFS with new functionality to fully support new
hash functions, one have to add a new capability (Zooko O'Whielacronx, personal
communication, September 29, 2010). This ensures backwards compatibility and can
be an optional feature for users to employ. However, we found that adding a new
capability was out of scope for this assignment.

% This chapter should contain (in listed and grouped order):
% -----------------------------------------------------------
% Describe inconsistencies in the numbers, and try to explain them.
% Time spent not hashing varies because: Variable use of CPU regs?
%
% Compare our results with general (SUPERCOP) results.
% Optimized candidates - which gain could be expected from more optimized
%   candidates?
%
% How much gain can Tahoe-LAFS get by switching to SHA-3?
%
% What if we had used another coding scheme? - does Cython lower the
%   candidates performance?
%
% More general stuff. What could have been done differently? Criticize
% ourselves.

@@

% This chapter should contain (in listed and grouped order):
% -----------------------------------------------------------
% A somewhat short conclusion summing up what has been done, and results and
% discussion
%
% We had hoped to have contributed more to the Open Source project that
% Tahoe-LAFS is.
